---
title: "Relationship between the Federal Funds Rate and 10 Year Treasury Yields"
author: "Sean Cannon"
date: "`r format(Sys.Date(),'%B %d, %Y')`"
output:
  html_document:
    df_print: paged
    code_folding: hide
    toc: no
    fig_caption: yes
    theme: sandstone
    toc_float: no
  pdf_document:
    toc: no
---
<style type="text/css">
h1.title {
  text-align: center;
  padding-bottom: 16px;
  padding-top: 16px;
}
h4.author {
  text-align: center;
}
h4.date {
  text-align: center;
}
</style>

---

```{r setup, include=FALSE}
rm(list=ls())
graphics.off()
# ECHO, MESSAGE, AND WARNING = FALSE for final knit
knitr::opts_chunk$set(echo = TRUE,
                      message = TRUE,
                      warning = TRUE)
```

```{r packages, message=FALSE, warning=FALSE}
require(fpp3)
require(tsibble)
require(tidyverse)
require(tidyquant)
require(kableExtra)
require(timetk)
require(vars)
require(lmtest)
```

<br>
<h3>Abstract</h3>
<br>

Write the abstract here. Include at the end: The report is also posted online [here](https://scannon2k.github.io/projects.html).


---


<br>
<h3>Introduction</h3>
<br>

Write the introduction here.


---


```{r data}
# Vector of variables to be gathered:
vars <- c("FEDFUNDS","GS10")

# Use tidyquant function tq_get to gather economic data from FRED and format as a tsibble
fred_data <- tq_get(vars,
                    get = "economic.data",
                    from = "1955-01-01") %>%
  mutate(Month = yearmonth(date), value = price) %>%
  select(-c(date, price)) %>%
  as_tsibble(index = Month, key = symbol)

# Pivot the data to be in a more conventional wide format
fred_dataw <- fred_data %>%
  pivot_wider(names_from = symbol, values_from = value) %>%
  as_tsibble()

# Use facet grid to visualize raw data for variables
fred_data %>%
  ggplot(aes(x = Month, y = value)) +
  geom_line() +
  facet_grid(vars(symbol), scales = "free_y") +
  labs(y = "Percent")

# Check for a unit root: Both variables have a unit root
# fred_data %>% 
#   features(value, unitroot_ndiffs) %>% 
#   kable(format = "html", table.attr = "style='width:30%;' ") %>% 
#   kableExtra::kable_styling()

# Transform variables
# tdata <- fred_dataw %>% select(c(FEDFUNDS,GS10)) %>%
#   mutate(infl = 1200*log(PCEPI/lag(PCEPI))) %>%                           # transformed inflation
#   mutate(dinfl = infl - lag(infl,1)) %>%                                  # differenced inflation
#   mutate(dinfl12 = 100*log(PCEPI/lag(PCEPI,12)) - lag(infl,12)) %>%       # differenced inflation 12
#   mutate(unrate = UNRATE - lag(UNRATE)) %>%                               # differenced unrate
#   mutate(expinf1yr = EXPINF1YR - lag(EXPINF1YR)) %>%                      # differenced expected inf
#   mutate(mich = MICH - lag(MICH)) %>%                                     # differenced mich
#   mutate(indpro = 1200*log(INDPRO/lag(INDPRO))) %>%                       # transformed indpro
#   # keep only transformed variables
#   select(-c(PCEPI, UNRATE, EXPINF1YR, MICH, INDPRO)) %>%
#   drop_na()

# Compare ACF plots before and after transformation
# fred_dataw %>% ACF(MICH) %>%
#   autoplot()
# tdata %>% ACF(mich) %>%
#   autoplot()

# Observe the plots of the transformed variables
# tdatam <- melt(tdata, "Month")
# ggplot(tdatam, aes(Month, value)) + 
#   geom_line() + 
#   facet_wrap(~variable, scales = "free", ncol = 2)
```

```{r}
Zdata <- fred_dataw %>%
  mutate(gFEDFUNDS = difference(log(FEDFUNDS)),
         gGS10 = difference(log(GS10))) %>%
  dplyr::select(gFEDFUNDS, gGS10) %>% 
  tidyr::drop_na()
zdata <- as.ts(Zdata, start = c(1955, 1), frequency = 12)

tmp <- VARselect(zdata,lag.max=12,type="const")
tmp
```

```{r}
p <- 1

var.rate <- VAR(zdata,p=p,type="const")
summary(var.rate)
```

```{r}
cat("H0: gFEDFUNDS does not cause gGS10,  H1: it does \n")
grangertest(gFEDFUNDS ~ gGS10,order=p,data=zdata)
```

```{r}
cat("H0: gGS10 does not cause gFEDFUNDS,  H1: it does \n")
grangertest(gGS10 ~ gFEDFUNDS,order=p,data=zdata)
```

```{r}
plot(irf(var.rate, n.ahead = 6, boot = TRUE))
```


<br>
<h3>Data and Variables</h3>
<br>

The data used in this analysis was gathered from the Federal Reserve’s online database (FRED), and the time series spans from January 1955 to March 2022. The table below shows a list of the variables used and their attributes.


```{r data_table}
# Define description and units vectors for the table
desc <- c("Federal Funds Effective Rate","Market Yield on U.S. Treasury Securities at 10-Year Constant Maturity")
units <- c("Percent, Not Seasonally Adjusted","Percent, Not Seasonally Adjusted")

# Create dataframe for kable function
table_data <- data.frame(vars, desc, units)

# Table displaying variables, description, and units
kbl(table_data, col.names = c("Variables", "Description", "Units"), align = "lll") %>%
  kable_styling(bootstrap_options = c("bordered", "hover"))
```
<br>

Write about the data and variables here.


---


```{r fit_models}
# Fit the models
```


<br>
<h3>Models</h3>
<br>


Talk about the models here.

<br>
<center>$\pi^{12}_{t}−\pi_{t−12}=\phi+\beta(B)\Delta\pi_{t−12}+\gamma(B)u_{t−12}+\varepsilon_t$</center>
<br>

Keep talking about the models here ($\bar{\pi}$).

<br>
<center>$me=(m1+m2+m3+m4)/4$</center>
<br>

---


```{r forecast_accuracy}
# Create the forecast using the combination model
```


<br>
<h3>Estimation and Results</h3>
<br>

Write about the results here.


```{r forecast_plot, fig.align='center'}
#Plot the estimated models

```
<br>

Keep writing about the results with latek maybe.

<br>
<center>$MAPE = mean(\left| p_t \right|)$</center>
<br>
<center>where, $p_t=100e_t/y_t$</center>


```{r insample_acc}
# Create a table to display in-sample accuracy between models

```
<br>

Results and tables etc.


```{r outsample_acc}
# Create a table to display out-of-sample accuracy between models

```
<br>

Closing findings.


---


<br>
<h3>Conclusion</h3>
<br>

Write about the conclusion here.


---


<br>
<h3>References</h3>
<br>
Hyndman, Rob J. “Forecasting: Principles&nbsp;and&nbsp;Practice (3rd Ed).” OTexts. Accessed April 10, 2022. https://otexts.com/fpp3/.
<br><br>
